---
title: "Full Analysis"
format: html
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(stringr)

combined_df <- read_excel("C:/Users/brody/Desktop/Brody Thesis Project/combined_coral_benthic_fish.xlsx")


# fish_df <- read_excel("C:/Users/brody/Desktop/Brody Thesis Project/Data Sets/Pal_Cage_fishvide_bm_v1.xlsx")
# 
# 
# cover_df <- read_excel("C:/Users/brody/Desktop/Brody Thesis Project/Data Sets/Pal_ExCag_Dat_BR_V1.xlsx") %>%
#   mutate(
#     Treatment = str_trim(Treatment),
#     Treatment = dplyr::recode(Treatment, "Control" = "Con")
#   )
```



```{r}
# library(readxl)
# library(dplyr)
# library(writexl)
# 
# # 1) Read + keep only the columns you need
# total_area_1 <- total_area %>%
#   select(Block, Treatment, Year, total_coral_area_cm2) %>%
#   mutate(Year = as.integer(Year) + 2017)
# 
# mean_indiv_growth_agg_1 <- mean_indiv_growth_agg %>%
#   select(Block, Treatment, Year, mean_indiv_growth_cm2, n_corals) %>%
#   mutate(Year = as.integer(Year) + 2017)
# 
# cover_df_1 <- cover_df %>%
#   select(Year, Block, Treatment,
#          CCA_calcareous, Fleshy_MA, `Lobo/Peys`, Scleractinian, SPAT, TAS) %>%
#   mutate(Year = as.integer(Year))
# 
# fish_df_1 <- fish_df %>%
#   select(Year, Block, Treatment, Grazerdetritivore, Herbivore_Sc_Br) %>%
#   mutate(Year = as.integer(Year))
# 
# # 2) Join everything together (keep all rows from total_area_norep)
# combined_df <- total_area_1 %>%
#   left_join(mean_indiv_growth_agg_1, by = c("Year", "Block", "Treatment")) %>%
#   left_join(cover_df_1, by = c("Year", "Block", "Treatment")) %>%
#   left_join(fish_df_1, by = c("Year", "Block", "Treatment")) %>%
#   select(
#     Year, Block, Treatment,
#     total_coral_area_cm2, mean_indiv_growth_cm2, n_corals,
#     CCA_calcareous, Fleshy_MA, `Lobo/Peys`, Scleractinian, SPAT, TAS,
#     Grazerdetritivore, Herbivore_Sc_Br
#   )
# 
# library(writexl)
# # 3) Export
# write_xlsx(combined_df, "C:/Users/brody/Desktop/Brody Thesis Project/Data Sets/combined_coral_benthic_fish.xlsx")
```

## Remove 2018 and 2024, convert year/treat/block to factor 

```{r}
library(dplyr)

combined_df1 <- combined_df %>%
  filter(!Year %in% c(2018, 2024)) |> 
   mutate(
    Year = factor(Year),
    Treatment = factor(Treatment),
    Block = factor(Block)
  )
```
## Reorder Treatments

```{r}
combined_df1 <- combined_df1 %>%
  mutate(
    Treatment = factor(
      Treatment,
      levels = c("Con", "1x1", "2x2", "TopOnly")
    )
  )
```



## Testing for Family and Model 

```{r}
check_response_1 <- function(n_corals){

  print(var)

  print(summary(combined_df1[[n_corals]]))

  print(mean(combined_df1[[n_corals]], na.rm=TRUE))
  print(var(combined_df1[[n_corals]], na.rm=TRUE))

  hist(combined_df1[[n_corals]])
  qqnorm(combined_df1[[n_corals]])
  qqline(combined_df1[[n_corals]])
}
```

```{r}
check_response_2 <- function(total_coral_area_cm2){

  print(var)

  print(summary(combined_df1[[total_coral_area_cm2]]))

  print(mean(combined_df1[[total_coral_area_cm2]], na.rm=TRUE))
  print(var(combined_df1[[total_coral_area_cm2]], na.rm=TRUE))

  hist(combined_df1[[total_coral_area_cm2]])
  qqnorm(combined_df1[[total_coral_area_cm2]])
  qqline(combined_df1[[total_coral_area_cm2]])
}
```

```{r}
check_response_3 <- function(mean_indiv_growth_cm2){

  print(var)

  print(summary(combined_df1[[mean_indiv_growth_cm2]]))

  print(mean(combined_df1[[mean_indiv_growth_cm2]], na.rm=TRUE))
  print(var(combined_df1[[mean_indiv_growth_cm2]], na.rm=TRUE))

  hist(combined_df1[[mean_indiv_growth_cm2]])
  qqnorm(combined_df1[[mean_indiv_growth_cm2]])
  qqline(combined_df1[[mean_indiv_growth_cm2]])
}
```


```{r}
check_response_2("total_coral_area_cm2")
check_response_3("mean_indiv_growth_cm2")
check_response_1("n_corals")
```

## Summary Stats of Models 

```{r}
summ_stats <- function(x) {
  c(
    n = sum(!is.na(x)),
    min = min(x, na.rm=TRUE),
    q1 = quantile(x, 0.25, na.rm=TRUE),
    median = median(x, na.rm=TRUE),
    mean = mean(x, na.rm=TRUE),
    q3 = quantile(x, 0.75, na.rm=TRUE),
    max = max(x, na.rm=TRUE),
    var = var(x, na.rm=TRUE)
  )
}

summ_stats(combined_df1$total_coral_area_cm2)
summ_stats(combined_df1$mean_indiv_growth_cm2)
summ_stats(combined_df1$n_corals)
```








## Rescale the data


```{r}
combined_df2 <- combined_df1 %>%
  mutate(across(
    c(CCA_calcareous, Fleshy_MA, `Lobo/Peys`,
      Scleractinian, SPAT, TAS,
      Grazerdetritivore, Herbivore_Sc_Br),
    scale
  ))
```

```{r}
library(dplyr)
library(corrplot)

# Select only numeric columns
numeric_df <- combined_df2 %>%
  select(where(is.numeric))

# Create correlation matrix
cor_matrix <- cor(numeric_df, use = "pairwise.complete.obs")

# Plot correlation matrix
corrplot(cor_matrix,
         method = "color",
         type = "upper",
         tl.col = "black",
         tl.srt = 45)
```












































```{r}
library(car)

full_model <- lm(
  mean_indiv_growth_cm2 ~
    Treatment +
    Year +
    CCA_calcareous +
    Fleshy_MA +
    `Lobo/Peys` +
    Scleractinian +
    SPAT +
    TAS +
    Grazerdetritivore +
    Herbivore_Sc_Br,
  data = combined_df2
)

vif(full_model)
```




## Creating the models before dredge

## Total area using gamma glm
```{r}
library(lme4)

full_area <- glmer(
  total_coral_area_cm2 ~ Treatment +
  Year + 
  CCA_calcareous +
  Fleshy_MA +
  `Lobo/Peys` +
  #Scleractinian +
  SPAT +
  TAS +
  #Grazerdetritivore +
  #Herbivore_Sc_Br +
  (1|Block),
  data = combined_df2,
  family = Gamma(link="log")
)
```


## Coral Growth using gaussian mixed model (simplest), maybe switch to gamma if residuals dont fit 

```{r}
full_growth <- lmer(
  mean_indiv_growth_cm2 ~ Treatment +
  Year +
  CCA_calcareous +
  Fleshy_MA +
  `Lobo/Peys` +
  Scleractinian +
  SPAT +
  TAS +
  Grazerdetritivore +
  Herbivore_Sc_Br +
  (1|Block),
  data = combined_df2
)
```


## Density of corals using generalized linear model with a negative binomial family 

```{r}
full_n <- glmer.nb(
  n_corals ~ Treatment +
  Year + 
  CCA_calcareous +
  Fleshy_MA +
  `Lobo/Peys` +
  Scleractinian +
  SPAT +
  TAS +
  Grazerdetritivore +
  Herbivore_Sc_Br +
  (1|Block),
  data = combined_df2
)
```


## Maybe need to scale variables and test for correlation for predictors 

## Dredge 

```{r, warning=FALSE, message=FALSE}
library(MuMIn)

options(na.action = "na.fail")

dredge_area <- dredge(full_area, fixed = "Year")
#dredge(full_growth)
#dredge(full_n)
#
#
best_area <- get.models(dredge_area, 1)[[1]]

summary(best_area)
```



```{r}
top_models <- get.models(dredge_area, subset = delta < 2)

avg_model <- model.avg(top_models)

summary(avg_model)
```

```{r}
sw(avg_model)
```

```{r}
best_area <- get.models(dredge_area, 1)[[1]]

library(DHARMa)

sim_res <- simulateResiduals(best_area)

plot(sim_res)
```








```{r, warning=FALSE, message=FALSE}
library(MuMIn)

options(na.action = "na.fail")

#dredge(full_area)
dredge(full_growth, fixed = "Year")
#dredge(full_n)
```




```{r}
top_models_growth <- get.models(dredge_growth, subset = delta < 2)

avg_model_growth <- model.avg(top_models_growth)

summary(avg_model_growth)
```

```{r}
sw(avg_model_growth)
```








```{r}
library(MuMIn)

options(na.action = "na.fail")

#dredge(full_area)
#dredge(full_growth)
dredge(full_n, fixed ="Year")
```